Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-03-26 document/ms.tex
more formatting

@@

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
  \emph{not} by weighting the detected objects by the inverse detection efficiency (CITE PETIGURA),
  but rather by forward modeling the observed distribution subject to the detection efficienty.
Re-weighting the data by inverse selection probability---%
  a method called V-max in the quasar and galaxy luminosity function contexts---%
  is not wrong,
  but it produces a higher-variance estimate of the population than a justified likelihood approach.
One simple intuition along these lines is that the inverse probability weights can be large in dynamic range,
  so the effective number of samples, after weighting, can be much smaller than the actual number of samples.
Another is that it tends to upweight most strongly the most marginal detections in the data sample,
  since these tend to be the samples that lie in the region of lowest completeness.
In this work, we build a justified likelihood of the observed data,
  taking into account the completeness function or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood;
  it makes the fundamental assumption that the data points are independently drawn from the model,
  but is agnostic about all other properties of the data or distribution functions.
Indeed, the assumptions behind this study are weaker than those of any previous study
  of the Earth-analog population (CITE PETIGURA, DONG).

@@

Perhaps the most important novelty of this study is in its treatment of observational uncertainties.
Although planet periods in our data set are very precisely measured,
  the planet radius measurements have large uncertainties.
The justified probabilistic approach to including these uncertainties in a distribution analysis
  is to introduce the ``true'' radii as latent variables and infer and marginalize them out
  along with all other nuisance parameters.
Although this sounds expensive---and it can be---%
  there is a simple approach we have been advocating (CITE HOGG) that makes use of importance sampling.
We use this approach here;
  our results are fully marginalized over posterior uncertainties in all of the planet radii.
More will be said about this below;
  the importance sampling looks like a product over data points
  of sums (averages) over individual-datum radius samplings.

>>>>

%
%  RULES OF THE GAME, MOTHAFUCKERS.
%
%  * 80 characters
%  * line breaks at the ends of sentences
%  * that is all.
%

@@

This study is novel in a number of ways.
We create probabilistic information about the full exoplanet population
\emph{not} by weighting the detected objects by the inverse detection
efficiency (CITE PETIGURA), but rather by forward modeling the observed
distribution subject to the detection efficienty.
Re-weighting the data by inverse selection probability---a method called
V-max in the quasar and galaxy luminosity function contexts---is not wrong,
but it produces a higher-variance estimate of the population than a justified
likelihood approach.
One simple intuition along these lines is that the inverse probability weights
can be large in dynamic range, so the effective number of samples, after
weighting, can be much smaller than the actual number of samples.
Another is that it tends to upweight most strongly the most marginal
detections in the data sample, since these tend to be the samples that lie in
the region of lowest completeness.  In this work, we build a justified
likelihood of the observed data, taking into account the completeness function
or detection efficiency.
Our likelihood function is a variable-rate Poisson likelihood; it makes the
fundamental assumption that the data points are independently drawn from the
model, but is agnostic about all other properties of the data or distribution
functions.
Indeed, the assumptions behind this study are weaker than those of any
previous study of the Earth-analog population (CITE PETIGURA, DONG).

@@

Perhaps the most important novelty of this study is in its treatment of
observational uncertainties.
Although planet periods in our data set are very precisely measured, the
planet radius measurements have large uncertainties.
The justified probabilistic approach to including these uncertainties in a
distribution analysis is to introduce the ``true'' radii as latent variables
and infer and marginalize them out along with all other nuisance parameters.
Although this sounds expensive---and it can be---there is a simple approach
we have been advocating (CITE HOGG) that makes use of importance sampling.
We use this approach here; our results are fully marginalized over posterior
uncertainties in all of the planet radii.
More will be said about this below; the importance sampling looks like a
product over data points of sums (averages) over individual-datum radius
samplings.
